Pipit on the Post: Proving Pre- and Post-conditions of Reactive Systems

### 2024-01-15 2024pipit/section/60-Evaluation.tex
clean up eval

We were able to verify the Kind2 implementation of the simplified trigger-fetch mechanism for trigger arrays up to size sixteen.
Verification of a 4-size array took 11s of CPU time; an 8-size array took 13s; a 16-size array took 276s; and a 32-size array timed out after \TODO{hours}.
In comparison, the Pipit version is verified \emph{for all sizes of arrays} and the checking took 5s of CPU time.
The M_TTCAN hardware implementation of TTCAN supports up to 64 triggers~\cite{bosch2019mttcan}.
Both developments of the simplified trigger-fetch logic are roughly the same size, on the order of two-hundred lines of code including comments.
We ran these experiments on a 2020 MacBook Air M1 with 16 gigabytes of RAM.

A critical simplification we made to the Kind2 implementation was to modify the trigger-enabled set to be a single cycle index.
In the specification, the enabled set is implemented as a cycle-offset and repeat-factor.
Checking if a trigger is enabled in the current cycle requires nonlinear arithmetic, which is difficult for SMT solvers.
In our Pipit development, we can treat the definition of the cycle set abstractly.
However, in the Kind2 development, quantifiers cannot contain function calls, which means that we cannot hide the implementation of the enabled-set check by providing an abstract contract.

% \begin{tabular}{rrr}
%   size & wall-clock & CPU time \\
%   4 & 8s & 11s \\
%   8 & 8s & 13s \\
%   16 & 125s & 276s \\
%   32 & - & - \\
% \end{tabular}

>>>>

We were able to verify the Kind2 implementation of the simplified trigger-fetch mechanism for trigger arrays containing up to 16 elements; above that, a 32-size array reported multiple runtime errors.
% Verification of a 4-size array took 11s of CPU time; an 8-size array took 13s; a 16-size array took 276s; and a 32-size array timed out after several hours and multiple runtime errors.
For reference, the M_TTCAN hardware implementation of TTCAN supports up to 64 triggers~\cite{bosch2019mttcan}.

We made a critical simplification in the Kind2 implementation, which was to modify the trigger-enabled set to be a single cycle index.
In the specification, the enabled set is implemented as a cycle-offset and repeat-factor.
Checking if a trigger is enabled in the current cycle requires nonlinear arithmetic, which is difficult for SMT solvers.
In our Pipit development, we can treat the definition of the cycle set abstractly.
However, in the Kind2 development, quantifiers cannot contain function calls, which means that we cannot hide the implementation of the enabled-set check by providing an abstract contract.

\autoref{f:evaluation:kind2-runtime} shows the verification runtime for different sizes of arrays; the Pipit version is parametric in the array size, and is thus verified for all sizes of arrays.
We ran these experiments on a 2020 M1 MacBook Air with 16 gigabytes of RAM.
Both Kind2 and Pipit developments of the simplified trigger-fetch logic are roughly the same size, on the order of two-hundred lines of code including comments.

\begin{figure}
  \center
\begin{tabular}{r|rr|rr|rr}
  & \multicolumn{4}{c|}{Kind2} & Pipit \\
  & \multicolumn{2}{c|}{simple enable-set} & \multicolumn{2}{c|}{full enable-set} & \\
  size & wall-clock & CPU & wall-clock & CPU & wall-clock & CPU \\
  \hline
  1 & 2s &  3s  & 6s & 12s & 6s & 6s \\
  2 & 3s &  3s  & 8s & 12s & 6s & 6s \\
  4 & 5s & 11s  & 12s & 24s & 6s & 6s \\
  8 & 8s & 13s  & 82s & 90s & 6s & 6s \\
  16 & 125s & 276s & \multicolumn{2}{c|}{(error)} & 6s & 6s \\
  32 & \multicolumn{2}{c|}{(error)} & \multicolumn{2}{c|}{(error)} & 6s & 6s \\
\end{tabular}
\caption{Verification time for trigger-fetch; simple enable-set uses a simplified version of the enable-set, while full enable-set uses bitwise arithmetic as in the TTCAN specification.}
\label{f:evaluation:kind2-runtime}
\end{figure}
